A load balancing scheme for massively multiplayer online games

### 2009-04-28 textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\begin{abstract}
MMOGs are applications which require much network bandwidth to function properly. In a distributed MMOG server architecture, with heterogeneous resources, the server nodes may become easily overloaded by the high demand from the players for state updates. Much work has been done in order to distribute the load on the server nodes, but this load is usually defined as the number of players on each server, what is not an ideal measure. Also, the heterogeneity of the system is frequently overlooked. In this work, we propose a balancing scheme which has two main goals: allocate load on the servers proportionally to the each one's power and reduce as much as possible the overhead from the distribution, considering the load as the occupied bandwidth of each server. It is is divided in three phases: local selection of servers, balancing and refinement. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers. We also make a review of related works and some comparisons were made.

@@

The main characteristic of massively multiplayer games is the large number of players, having dozens, or even hundreds, of thousands of participants simultaneously \cite{schiele2007rpp}. This large number of players interacting with one another generates a traffic on the support network which may grow quadratically compared to the number of players, in the worst case (Figure \ref{fig:quadratic}).

@@

It is important to notice that the while condition (line \ref{alg:ggp:whilefraction}) may lead to the assignment of a cell whose weight is higher than the weight share left on the region. However, it is very unlikely to assign to a region its exact weight share, so it is preferrable to surpass this value, guarateeing that every cell will be assigned to some region. Also, assigning to a region a weight which is higher than its calculated share does not imply that it will become overloaded. First, the weight share of a region is not equal to the power of its server. When there are enough resources, the region's weight share will be smaller than its capacity. Furthermore, as \ggp\ is a greedy algorithm, the heaviest cells will be assigned first, to the most powerful servers.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for massively multiplayer online games %\thanks{}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\begin{abstract}
MMOGs, or massively multiplayer online games, are applications which require much network bandwidth to function properly. In a distributed MMOG server architecture, with heterogeneous resources, the server nodes may become easily overloaded by the high demand from the players for state updates. Much work has been done in order to distribute the load on the server nodes, but this load is usually defined as the number of players on each server, what is not an ideal measure. Also, the heterogeneity of the system is frequently overlooked. In this work, we propose a balancing scheme which has two main goals: allocate load on the servers proportionally to the each one's power and reduce as much as possible the overhead from the distribution, considering the load as the occupied bandwidth of each server. It is divided in three phases: local selection of servers, balancing and refinement. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers. We also make a review of related works and some comparisons were made, where our approach performed better.

@@

The main characteristic of massively multiplayer online games is the large number of players, having dozens, or even hundreds, of thousands of participants simultaneously. This large number of players interacting with one another generates a traffic on the support network which may grow quadratically compared to the number of players \cite{chen2006gta}, in the worst case (Figure \ref{fig:quadratic}).

@@

It is important to notice that the while condition (line \ref{alg:ggp:whilefraction}) may lead to the assignment of a cell whose weight is higher than the weight share left on the region. However, it is very unlikely to assign to a region its exact weight share, so it is preferable to surpass this value, guaranteeing that every cell will be assigned to some region. Also, assigning to a region a weight which is higher than its calculated share does not imply that it will become overloaded. First, the weight share of a region is not equal to the power of its server. When there are enough resources, the region's weight share will be smaller than its capacity. Furthermore, as \ggp\ is a greedy algorithm, the heaviest cells will be assigned first, to the most powerful servers.

@@

\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
	\includegraphics[width=0.2\linewidth]{images/Fig-Bezerra.jpg}

	\textbf{Carlos Eduardo Benevides Bezerra} is a full-time Ph.D. student at UFRGS (Universidade Federal do Rio Grande do Sul). He obtained his M.Sc. from the same university in 2009 and his B.Sc. degree from UFBA (Universidade Federal da Bahia) in 2007.
%% figure caption is below the figure
%\caption{Please write your figure caption here}
\label{fig:cebb}       % Give a unique label
\end{figure}

\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
  \includegraphics[width=0.2\linewidth]{images/Fig-Geyer.jpg}

  \textbf{Cláudio Fernando Resin Geyer} is a full-time Professor at UFRGS.
%% figure caption is below the figure
%\caption{Please write your figure caption here}
\label{fig:cebb}       % Give a unique label
\end{figure}
